Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-22 talk/pepm2011/paper.tex
fix citation
kill XXX, remove line overflow

The method calls in the trace are always preceded by a \lstinline{guard_class}
operation, to check that the class of the receiver is the same as the one that
was observed during tracing.\footnote{\lstinline{guard_class} XXX lstinline too large in footnotes
performs a precise
class check, not checking for subclasses.} These guards make the trace specific
to the situation where \lstinline{y} is really a \lstinline{BoxedInteger}. When
the trace is turned into machine code and afterwards executed with
\lstinline{BoxedFloat}, the
first \lstinline{guard_class} instruction will fail and execution will continue
using the interpreter.

@@

In this section we want to give a formal description of the semantics of the
traces and of the optimizer and liken the optimization to partial evaluation.
We focus on the operations for manipulating heap allocated objects,
as those are the only ones that are actually optimized. We also consider only
objects with two fields $L$ and $R$ in this section, generalizing to arbitrary many fields
is straightforward.

Traces are lists of operations. The operations considered here are
\lstinline{new}, \lstinline{get}, \lstinline{set} and \lstinline{guard_class}.
The values of all
variables are locations (\ie pointers). Locations are mapped to objects, which
are represented by triples $(T,l_1,l_2)$ of a type $T$, and two locations that represent the
fields of the object. When a new object is created, the fields are initialized
to null, but we require that they are initialized to a real
location before being read, otherwise the trace is malformed (this condition is
guaranteed by how the traces are generated in PyPy).

@@

@@

Partially known data structures are built directly into Prolog (via unbound
logic variables) and thus the treatment of partially static data structures
was part of partial evaluation of Prolog programs from the early stages
\cite{lloyd_partial_1991}. One effect of unfolding in Prolog is that terms that
are constructed and immediately matched again, completely disappear in the
residual program. This is similar to what our optimization does for an
imperative language.
In functional programming this idea was introduced as
 constructor specialisation by Mogensen \cite{Mogensen:PEPM93}.

>>>>

Method calls in the trace are preceded by a \lstinline{guard_class}
operation, to check that the class of the receiver is the same as the one that
was observed during tracing.\footnote{\lstinline{guard_class}
performs a precise
class check, not checking for subclasses.} These guards make the trace specific
to the situation where \lstinline{y} is really a \lstinline{BoxedInteger}. When
the trace is turned into machine code and afterwards executed with
\lstinline{BoxedFloat}, the
first \lstinline{guard_class} instruction will fail and execution will continue
using the interpreter.

@@

In this section we want to give a formal description of the semantics of the
traces and of the optimizer and liken the optimization to partial evaluation.
We focus on the operations for manipulating heap allocated objects,
as those are the only ones that are actually optimized. We also consider only
objects with two fields $L$ and $R$ in this section, generalizing to arbitrary many fields
is straightforward. Traces are lists of operations. The operations considered here are
\lstinline{new}, \lstinline{get}, \lstinline{set} and \lstinline{guard_class}.

The values of all
variables are locations (\ie pointers). Locations are mapped to objects, which
are represented by triples $(T,l_1,l_2)$ of a type $T$, and two locations that represent the
fields of the object. When a new object is created, the fields are initialized
to null, but we require that they are initialized to a real
location before being read, otherwise the trace is malformed (this condition is
guaranteed by how the traces are generated in PyPy).

@@

% subsection Operational Semantics of Traces (end)

@@

Partially known data structures are built directly into Prolog (via unbound
logic variables) and thus the treatment of partially static data structures
was part of partial evaluation of Prolog programs from the early stages
\cite{lloyd_partial_1991}. One effect of unfolding in Prolog is that terms that
are constructed and immediately matched again, completely disappear in the
residual program. This is similar to what our optimization does for an
imperative language.
In functional programming this idea was introduced as
 constructor specialisation by Mogensen \cite{mogensen_constructor_1993}.
